---
title: "Lab 07 - Web scraping and Regular Expressions-pm566"
author: "Xiaoyu Zhu"
date: "10/8/2021"
output:
  html_document: 
  github_document: 
    html_preview: false
  word_document: default
always_allow_html: true
---


```{r setup, include=FALSE}
## packages
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
options(repos = c(CRAN = "http://cran.rstudio.com"))
library(tidytext)
library(tidyverse)
library(httr)
library(xml2)
library(stringr)
```

# Lab description
In this lab, we will be working with the NCBI API to make queries and extract information using XML and regular expressions. For this lab, we will be using the httr, xml2, and stringr R packages.

This markdown document should be rendered using github_document document.


# Question 1: How many sars-cov-2 papers?
Build an automatic counter of sars-cov-2 papers using PubMed. You will need to apply XPath as we did during the lecture to extract the number of results returned by PubMed in the following web address:
https://pubmed.ncbi.nlm.nih.gov/?term=sars-cov-2

```{r }
if (knitr::is_html_output(excludes = "gfm")){}
```


```{r}
# Downloading the website
website <- xml2::read_html("https://pubmed.ncbi.nlm.nih.gov/?term=sars-cov-2")
# Finding the counts
counts <- xml2::xml_find_first(website, "/html/body/main/div[9]/div[2]/div[2]/div[1]/div[1]")
# Turning it into text
counts <- as.character(counts)
# Extracting the data using regex
stringr::str_extract(counts, "[0-9,]+")


stringr::str_extract(counts, "[[:digit:],]+")
stringr::str_replace(counts, "[^[:digit:]]+([[:digit:]]+),([[:digit:]]+)[^[:digit:]]+", "\\1\\2")
```


# Question 2: Academic publications on COVID19 and Hawaii
```{r}
library(httr)
query_ids <- GET(
  url   = "https://eutils.ncbi.nlm.nih.gov/entrez/eutils/esearch.fcgi",
  query = list(
    db="pubmed",
    term="covid19 hawaii",
    retmax=1000)
)

# Extracting the content of the response of GET
ids <- httr::content(query_ids)
ids
```



# Question 3: Get details about the articles
```{r analyzing}
# Turn the result into a character vector
ids <- as.character(ids)
# Find all the ids 
ids <- stringr::str_extract_all(ids, "<Id>[[:digit:]]+</Id>")[[1]]
# Remove all the leading and trailing <Id> </Id>. Make use of "|"
# stringr::str_remove_all(ids, "</?Id>")
ids <- stringr::str_remove_all(ids, "<Id>|</Id>")
head(ids)
```

```{r retrieving-papers}
publications <- GET(
  url   = "https://eutils.ncbi.nlm.nih.gov/",
  path  = "entrez/eutils/efetch.fcgi",
  query = list(
    db = "pubmed",
    id = I(paste(ids, collapse=",")),
    retmax = 1000,
    rettype = "abstract"
    )
)
# Turning the output into character vector
publications <- httr::content(publications)
publications_txt <- as.character(publications)
```

# Question 4: Distribution of universities, schools, and departments
Using the function stringr::str_extract_all() applied on publications_txt, capture all the terms of the form:

University of …
… Institute of …
Write a regular expression that captures all such instances
```{r}
library(stringr)
institution <- stringr::str_extract_all(
  str_to_lower(publications_txt),
  "[[:alpha:]-]+university|university\\s+of\\s+(southern|new|northern|the)?\\s*[[:alpha:]-]+|[[:alpha:]-]+\\s+institute\\s+of\\s+[[:alpha:]-]+"
  ) 

# institution <- unlist(institution)
table(institution)
```

Repeat the exercise and this time focus on schools and departments in the form of

School of …
Department of …
And tabulate the results
```{r}
schools_and_deps <- str_extract_all(
  str_to_lower(publications_txt),
  "[[:alpha:]-]+\\s+school\\s+of\\s+(the)?\\s*[[:alpha:]-]+\\s|[[:alpha:]-]+\\s+department\\s+of\\s+[[:alpha:]-]+\\s"
  )
table(schools_and_deps)
```




# Question 5: Form a database
We want to build a dataset which includes the title and the abstract of the paper. The title of all records is enclosed by the HTML tag ArticleTitle, and the abstract by Abstract.

Before applying the functions to extract text directly, it will help to process the XML a bit. We will use the xml2::xml_children() function to keep one element per id. This way, if a paper is missing the abstract, or something else, we will be able to properly match PUBMED IDS with their corresponding records.


```{r}

pub_char_list <- xml2::xml_children(publications)
pub_char_list <- sapply(pub_char_list, as.character)
```

```{r extracting-last-bit, eval = FALSE}
abstracts <- stringr::str_extract(pub_char_list, "<Abstract>(\\n|.)+</Abstract>")
abstracts <- stringr::str_remove_all(abstracts, "</?[[:alnum:]]+>")
abstracts <- stringr::str_replace_all(abstracts, "\\s+", " ")
table(is.na(abstracts))
```


How many of these don't have an abstract? Answer: There are 9 articles without abstracts.

Now, the title

```{r process-titles, eval = FALSE}
titles <- str_extract(pub_char_list, "<ArticleTitle>(\\n|.)+</ArticleTitle>")
titles <- str_remove_all(titles, "</?[[:alnum:]]+>")
titles <- str_replace_all(titles, "\\s+", " ")
table(is.na(titles))
```

There are no articles without titles.

Finally, put everything together into a single `data.frame` and use
`knitr::kable` to print the results

```{r build-db, eval = FALSE}
database <- data.frame(
  PubMedID  = ids,
  Title     = titles,
  Abstracts = abstracts
)
knitr::kable(database)
```

Done! Knit the document, commit, and push.

## Final Pro Tip (optional)

You can still share the HTML document on github. You can include a link in your `README.md` file as the following:

```md
View [here](https://ghcdn.rawgit.org/:user/:repo/:tag/:file)
```

For example, if we wanted to add a direct link the HTML page of lecture 7, we could do something like the following:

```md
View week7-lab[here](https://ghcdn.rawgit.org/lysethan/PM566-labs/master/week7/week7-lab.html
```


