---
title: 'Lab 10 - SQL'
author: "Xiaoyu Zhu"
date: "11/5/2021"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(RSQLite)
library(DBI)
library(sqldf)
```

```{r}
# Initialize a temporary in memory database
con <- dbConnect(SQLite(), ":memory:")

# Download tables
actor <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/actor.csv")
rental <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/rental.csv")
customer <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/customer.csv")
payment <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/payment_p2007_01.csv")

# Copy data.frames to database
dbWriteTable(con, "actor", actor)
dbWriteTable(con, "rental", rental)
dbWriteTable(con, "customer", customer)
dbWriteTable(con, "payment", payment)
```


```{r}
dbListTables(con)
```


IP: Use can use the following QUERY to see the structure of a table

You can also use knitr + SQL!

```{sql, connection=con, output.var="x1"}
PRAGMA table_info(actor)
```

```{r}
x1
```

This is equivalent to use `dbGetQuery`

```{r}
dbGetQuery(con, "PRAGMA table_info(actor)")
```

# Exercise 1
### Retrive the actor ID, first name and last name for all actors using the actor table. Sort by last name and then by first name.

```{r}
dbGetQuery(con, "
/* This is COMMENT! */
SELECT actor_id, first_name, last_name
FROM actor /* YOU CAN ADD COMMENTS USING
MULTIPLE LINES! */
ORDER by last_name, first_name 
LIMIT 5")

```

# Exercise 2
### Retrive the actor ID, first name, and last name for actors whose last name equals ‘WILLIAMS’ or ‘DAVIS’.

```{r}
dbGetQuery(con, "
/* This is COMMENT! */
SELECT actor_id, first_name, last_name
FROM actor /* YOU CAN ADD COMMENTS USING
MULTIPLE LINES! */
WHERE last_name IN ('WILLIAMS', 'DAVIS')")
```

# Exercise 3
## Write a query against the rental table that returns the IDs of the customers who rented a film on July 5, 2005 (use the rental.rental_date column, and you can use the date() function to ignore the time component). Include a single row for each distinct customer ID.

```{r}
dbGetQuery(con, "PRAGMA table_info(rental)")
```
```{r}
dbGetQuery(con," 
SELECT DISTINCT customer_id 
FROM rental
WHERE date(rental_date) = '2005-07-05' LIMIT 5")
```


# Exercise 4
## Exercise 4.1
### Construct a query that retrives all rows from the payment table where the amount is either 1.99, 7.99, 9.99.
```{r}
q <- dbSendQuery(con, "
SELECT *
FROM payment
WHERE amount IN (1.99, 7.99, 9.99)"
)
dbFetch(q, n = 10)
dbFetch(q, n = 10)
dbClearResult(q)
```



## Exercise 4.2
### Construct a query that retrives all rows from the payment table where the amount is greater then 5

```{r}
dbGetQuery(con, "
SELECT *
FROM payment
WHERE amount > 5 LIMIT 5")
```

## Exercise 4.3
### Construct a query that retrives all rows from the payment table where the amount is greater then 5 and less then 8

SELECT *
FROM ___
WHERE ___ AND ___
```{r}
dbGetQuery(con,"
SELECT staff_id, COUNT(*) AS N
FROM payment
WHERE amount>5 AND amount<8
GROUP BY staff_id 
")
```


# Exercise 5
### Retrive all the payment IDs and their amount from the customers whose last name is ‘DAVIS’.


```{r}
dbGetQuery(con, "
SELECT p.payment_id, p.amount
FROM payment AS p
  INNER JOIN customer AS c ON p.customer_id = c.customer_id
WHERE c.last_name = 'DAVIS'")
```

# Exercise 6
## Exercise 6.1
### Use COUNT(*) to count the number of rows in rental
```{r}
dbGetQuery(con,"
SELECT rental_id, 
COUNT(*) AS nrows
FROM rental
")
```




## Exercise 6.2
### Use COUNT(*) and GROUP BY to count the number of rentals for each customer_id

```{r}
dbGetQuery(con,"
SELECT customer_id, COUNT(*) AS 'number of each customer id'
FROM rental
GROUP BY customer_id
LIMIT 10
")
```

## Exercise 6.3
### Repeat the previous query and sort by the count in descending order
```{r}
dbGetQuery(con,"
SELECT customer_id, 
COUNT(*) AS 'number of each customer id'
FROM rental
GROUP BY customer_id
/*
This is equivalent to
ORDER BY -`N Rentals` LIMIT 5
*/
ORDER BY 'number of each customer id' DESC
LIMIT 10
")
```

## Exercise 6.4
### Repeat the previous query but use HAVING to only keep the groups with 40 or more.

```{r}
dbGetQuery(con,"
SELECT COUNT(*) AS num_rentals, customer_id
FROM rental
GROUP BY customer_id
HAVING num_rentals >= 40
ORDER BY num_rentals DESC
")
```


# Exercise 7
### The following query calculates a number of summary statistics for the payment table using MAX, MIN, AVG and SUM
```{r}
dbGetQuery(con,"
SELECT MAX(amount) as 'max of payment',
       MIN(amount) as 'min of payment',
       AVG(amount) as 'avg of payment',
       SUM(amount) as 'sum of payment',
       customer_id
FROM payment
LIMIT 5
")
```

## Exercise 7.1
### Modify the above query to do those calculations for each customer_id

```{r}
dbGetQuery(con,"
SELECT MAX(amount) as 'max of payment',
       MIN(amount) as 'min of payment',
       AVG(amount) as 'avg of payment',
       SUM(amount) as 'sum of payment',
       customer_id
FROM payment
GROUP BY customer_id
LIMIT 10
")
```

## Exercise 7.2
### Modify the above query to only keep the customer_ids that have more then 5 payments
```{r}
dbGetQuery(con,"
SELECT MAX(amount) as max,
       MIN(amount) as min,
       AVG(amount) as avg,
       SUM(amount) as sum,
       COUNT(*) as count,
       customer_id
FROM payment
GROUP BY customer_id
HAVING COUNT(*)>5
")
```


# Cleanup
### Run the following chunk to disconnect from the connection.

```{r cleaning}
dbDisconnect(con)
```

